---
title: "ReadMe for ntbgraphics"
last update: 02/25/2020
output:
  rmarkdown::github_document:
  toc: true
---

# Table of contents
* [Introduction]
* [Basic Principles]
* [Examples]
* [Installation]
* [Demo]


# Introduction 

The following introduction to the R package 'ntbgraphics' aims not only for expounding its features but also for facilitating access to it, even for relatively unexperienced R users. Please refer to the [Installation] part if you have not yet downloaded R itself.

This package includes functions for importing, transforming and visualization of NTB datasets:

  - **'getexpdata'** for import of the Animal List (containing animal *RFIDs*, corresponding Genotypes and optionally Environment) and Meta Behavior (containing columns *Animal* and the animal's behavioral measures) as well as formal preparation;
  - **'ploteachexp'** for plotting of given experiments as boxplots and exporting the result as a PDF file;
  - **'loopplotexp'** for plotting all experiments within a dataset as boxplots and exporting the results in one PDF file;
  - **'heatmapexp'** for plotting all experiments as a heatmap and producing a datamatrix with z-scored values;
  - **'pcatsneexp'** for PCA and tSNE results and cluster plots;
  - **'colordiverger'** for creating a customized diverging color palette for visualization in different contexts.
  
[Go to Top](#top)
  
# Basic Principles

All functions (apart from 'colordiverger') take a **directory** as their input, which specifies the location of the **two files** "Animal List.xlsx" and "Meta Behavior.xlsx" (mind correct spelling of these files - functions rely on specific names as given!). Furthermore, the user has to specify the kind of data provided, respectively the kind of analysis he wants to perform by defining the correct **analysis**. Please refer to the 'getexpdata' documentation for further information! 

Alternativley, you might call the arguments each function takes by executing the following line of code:

``` {r eval=FALSE}

## you may look for every other function the same way
formals(getexpdata)

## note: in case of many options for an argument, the first one listed also is the default setting
 
```

It is also important that you mind correct **formatting** of your excel files. This includes:

  - at least two columns with information about 'RFID' and 'Genotype'/'Environmental'/'Treatment' in your Animal List with these exact titles;
  - at least one column with information about 'Animal' (matching the information in the 'RFID' column in the Animal List), and at least one behavioral test in your Meta Behavior with exact titles:
\   
 _"Animal" "Meanspeed" "Rotations" "Center"	"Alternations" "Choices" "Context" "Cue" "FreezeBase" "Timeimmobile" "Baseline"	"inhibition70" "inhibition75" "inhibition80" "SucPref" "PlacePref" "ReversalLearn" "Activity" "Nocturnal" "SerialLearn"_
\
  _(-> this is the current entity of all available experiment names for analysis and plotting/mapping; if you need to add more experiments to this list, please refer to the creator of this package)_
\   

Further aspects can be customized depending on the specific function within that function.

All functions externally work on their own, which means that they may rely internally on one of the other functions of the package without the user needing to run them in advance.  

For heatmapping and especially for PCA and tSNE, it is important to have a sufficient number of animals for each group that ran through all of the tests that have been performed in that cohort or that will be analyzed. 
While 'heatmapexp' will display NAs (missing values) as 0 from the z-scored matrix and therefor simply in a neutral color - which will still lead to an interpretable result -, PCA and tSNE will exclude all animals with missing values completely, and thus might look quite poor with low animal numbers!

[Go to Top](#top)
 
# Examples

In this section you can find some examples how the output of some of the functions could look like. Be aware of the fact that you can customize many aspects of these outputs; refer to the functions' help pages for further information.
\   

The following plot shows the general layout you can expect from the boxplot functions (**'ploteachexp'** and **'loopplotexp'**)(please note: random data; therefore most likely no convincing differences between groups).

``` {r echo=FALSE, message=FALSE, warning=FALSE}

library(ntbgraphics)

ploteachexp(directory = paste0(system.file("extdata", package = "ntbgraphics",
                                                                mustWork = T),"/"),
            expname = "Meanspeed",
            analysis = "4arm_sd_tg",
            orderlevelcond = "gtblock",
            saveplotdir = FALSE)
```
\
\
\   

The following map shows the general layout you can expect from the **'heatmapexp'** function (please note: random data; therefore most likely no convincing clustering).

``` {r echo=FALSE, message=FALSE, warning=FALSE}

a <- heatmapexp(directory = paste0(system.file("extdata", package = "ntbgraphics",
                                                                mustWork = T),"/"),
                analysis = "4arm_sd_tg",
                orderlevelcond = "gtblock",
                acceptable.nas = 0,
                cutree_rows = 4,
                palette = "cRP",
                title = "Test Data Heatmap",
                saveplotdir = FALSE)
```
\
\
\   

The following plots show the general layout you can expect from the **'pcatsneexp'** function (please note: random data; therefore most likely no convincing clustering).

``` {r echo=FALSE, message=FALSE, warning=FALSE}

b <- pcatsneexp(directory = paste0(system.file("extdata/", package = "ntbgraphics", mustWork = T),"/"),
                analysis = "4arm_sd_tg",
                orderlevelcond = "gtblock",
                perplex =  10,
                theta = 0.8,
                ellipse_tsne = TRUE,
                ellconf = 0.3, 
                pastetitle = "Test PCA",
                pastetitle2 = "Test tSNE",
                saveplotdir = FALSE)
```
\
[Go to Top](#top)
\
\
\ 


# Installation

If you want to install this package, the following lines of code provide a simple way that does not rely on any dependencies except from you **having installed [R from CRAN](https://cran.r-project.org/mirrors.html) and ideally a GUI of your choice, e.g. [RStudio](https://rstudio.com/products/rstudio/)**. You may then copy all of the lines in the following paragraph or skip certain lines if redundant. For further information regarding usage of R, check out the following [introduction](https://cran.r-project.org/doc/contrib/Paradis-rdebuts_en.pdf) or any other introduction you may find useful.

``` {r eval=FALSE}

## install package devtools to get - amongst others - functions to access ntbgraphics 
## (and every other package on GitHub)
install.packages("devtools")
## install ntbgraphics with function 'install_github' from package devtools
devtools::install_github("volkmannp/ntbgraphics")
## that is it!

## you are now ready to use ntbgraphics on your computer and may load it using...
library(ntbgraphics)

## for examples on how to use the functions, read on below

```

For your information: You can use the command 'install.packages' for the installation of devtools due to its availability on CRAN (Comprehensive R Archive Network). Since ntbgraphics is not part of CRAN, its installation needs another function ('install_github') that in turn is part of the devtools package.

[Go to Top](#top)

# Demo

After installing ntbgraphics, you probably want to explore the package with some random data or might simply be curious how to specifically deal with the functions provided. Thus, below you may find some lines of code that address this inquisitiveness. However, they do not explore every single possible option available for the functions of this package which you may try out on your own.
\   

The example data used for the following is provided within the package by being included in the installed files.
You may simply copy all lines and run them at once or copy indivdual lines/functions you have a particular interest in.
\   

Note: Each function works independently of what you may have run in advance as long as 'ntbgraphics' has been loaded. Although the example shows the 'getexpdata' function as its very first, running it is not necessary for the other functions to work. This holds true for every single function!
\   
Also note that the directory within these lines aims for working on every computer by accessing the data provided within this package. If you want to use your own files, the directory might rather look like this: "/Users/specificuser/Documents/experiments/ntb/run1"

``` {r eval=FALSE}

## clear workspace and load package (with all functions and necessary dependencies included automatically)
rm(list = ls(all.names = TRUE))
library(ntbgraphics)

## (getexpdata) get modified table with data
data.animal.joined <- getexpdata(directory = paste0(system.file("extdata/", package = "ntbgraphics", 
                                                                mustWork = T),"/"),
                                 analysis = "4arm_sd_tg")

## (ploteachexp) plot a defined experiment
ploteachexp(expname = "Meanspeed",
            directory = paste0(system.file("extdata", package = "ntbgraphics", mustWork = T),"/"),
            analysis = "4arm_sd_tg",
            saveplotdir = FALSE)

## (loopplotexp) plot all experiments
loopplotexp(directory = paste0(system.file("extdata", package = "ntbgraphics", mustWork = T),"/"),
            analysis = "4arm_sd_tg")

## (heatmapexp) print out heatmap
data.animal.matrix <- heatmapexp(directory = paste0(system.file("extdata", package = "ntbgraphics",
                                                                mustWork = T),"/"),
                                 analysis = "4arm_sd_tg",
                                 saveplotdir = FALSE)

## (pcatsneexp) plot PCA and tSNE
results <- pcatsneexp(directory = paste0(system.file("extdata/", package = "ntbgraphics", 
                                                     mustWork = T),"/"),
                      analysis = "4arm_sd_tg",
                      perplex =  7,
                      saveplotdir = FALSE)
### -> access results of pcatsneexp (requires to run pcatsneexp and store results as shown above)
results_pca <- results[["pca_analysis"]]
results_tsne <- results[["tsne_analysis"]]

```

[Go to Top](#top)